Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-08-05 Paper/method.tex
..

In order to motivate the iterative strategy used in this work, we implement Algorithm~\ref{alg:grad} to the
19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
verification, $\hat{\mat{C}}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
the comparison of the resulting eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
relative L-2 norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
individual components of the first five eigenvectors of $\mat{C}$, constructed using $N$~=~1000 samples,
and using a smaller sample size, $n$ ranging from 20 to 240.} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
Interestingly, we observe that the dominant eigenvalues ($\lambda_1 - \lambda_4$) are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvalues. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we explore the applicability of a so-called gradient-free approach as mentioned earlier, in the
following section.

>>>>

In order to motivate the iterative strategy used in this work, we implement Algorithm~\ref{alg:grad} to the
19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
verification, $\hat{\mat{C}}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
the comparison of the resulting eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
relative L-2 norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
Interestingly, we observe that the dominant eigenvalues ($\lambda_1 - \lambda_4$) are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvectors. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we explore the applicability of a so-called gradient-free approach as mentioned earlier, in the
following section.
